### 2020-11-26 pilot_data_analysis.Rmd

---
title: "Pilot data analysis"
output:
  html_document:
    df_print: paged
---

```{r}
library(lmerTest)
library(tidyverse)
library(cowplot)
library(stargazer)
source("RainCloudPlots.R")

csv_data <- fs::dir_ls(path = "./data", glob = "*.csv") %>%
  map_dfr(read_csv, .id = "source", col_type = cols(
  .default = col_character(), rt = col_double(), corr = col_integer(), TrialID = col_integer())) %>% 
  filter(!is.na(TrialID) & TrialID < 1000) %>%
  select(source, participant, date, OS, frameRate, rt, corr, TrialID, StimulusType, Condition, PrimeDuration, Prime, Target)

# get participants with correct responses

accuracy_by_participant <- csv_data %>% group_by(source) %>% summarise(acc = mean(corr == 1), N = n())

participants_to_include <- accuracy_by_participant %>% filter(N == 480 & acc > .8)

participant_list_for_prolific <- str_extract(string = participants_to_include$source, pattern = "[a-z0-9]{24}")
paste0(participant_list_for_prolific, sep = "\n") %>% writeClipboard() # works on windows

# for now, we only have two participants with accuracy > .8 and a full data set

data_to_include <- csv_data %>% filter(source %in% participants_to_include$source) %>% mutate(StimulusType = StimulusType %>% factor, Condition = Condition %>% factor, PrimeDuration = PrimeDuration %>% factor, rt = rt * 1000)

write_csv(data_to_include, path = "current_data.csv")

data_summary_rt <- data_to_include %>% group_by(StimulusType, Condition, PrimeDuration) %>% summarise(M = mean(rt), SD = sd(rt), N = n())

# data above 1500 ms excluded

plot_raincloud(data_to_include %>% filter(corr == 1 & rt < 1500), x_column = "Condition", y_column = "rt", x_label = "Condition", y_label = "RT in ms (correct responses)", main_title = "RT by condition and word type") + 
  # put labels on the mean for easier comparison
  stat_summary(fun.data = function(x) {data.frame(label = format(mean(x), digits = 2), y = mean(x))}, 
                                      geom = "text",
                                      #width = 1,
                                      position = position_nudge(x = 0.5, y = 0),
                                      #size = 0.2,
                                      color = "black") +
  facet_grid(cols = vars(PrimeDuration), rows = vars(StimulusType), labeller = label_context)

```
## LMM for RT (words, correct responses)
```{r lmms}
lmm1 <- lmer(data = data_to_include %>% filter(corr == 1 & StimulusType == "WORD" & rt < 1500), rt ~ Condition * PrimeDuration + (1|source) + (1|Target))

summary(lmm1)

```
gLMM for accuracy (Words)
```{r logistic_lmm}
lmm_acc <- glmer(data = data_to_include %>% filter(corr != -1 & StimulusType == "WORD" & rt < 1500), corr ~ Condition * PrimeDuration + (1|Target), family = binomial(link = "logit"))

summary(lmm_acc)

```

Creepy: visualize participant locations from Qualtrics
```{r participant_map}
library(sp)
library(leaflet)

participant_data <- read_csv("participant_data.csv")
coordinates(participant_data) <- c("Location Longitude","Location Latitude")
leaflet(participant_data) %>% addMarkers(label = .$`Confirm Prolific ID`) %>% addTiles()
```