### 2021-01-04 Manuscript.Rmd
Make manuscript files

---
title             : "Effects of masked prime duration in an online lexical decision task"
shorttitle        : "Online lexical decision task"

author: 
  - name          : "Bernhard Angele"
    affiliation   : "1,2"
    corresponding : yes    # Define only one corresponding author
    address       : "Department of Psychology, Faculty of Science and Technology, Talbot Campus, Fern Barrow, Poole BH12 5BB, UK"
    email         : "bangele@bournemouth.ac.uk"
  - name          : "Ana Baciero"
    affiliation   : "2,3"
  - name          : "Pablo Gomez"
    affiliation   : "3,4"
  - name          : "Manuel Perea Lara"
    affiliation   : "2,3"

affiliation:
  - id            : "1"
    institution   : "Bournemouth University"
  - id            : "2"
    institution   : Universitat de Valencia, Valencia, Spain
  - id            : "3"
    institution   : "Nebrija University, Madrid, Spain"
  - id            : "4"
    institution   : "California State University San Bernardino, Palm Desert Campus"

authornote: |
  Add complete departmental affiliations for each author here. Each new line herein must be indented, like this line.

  Enter author note here.

abstract: |
  Masked priming is one of the most important paradigms in the study of visual word recognition and representation.

  Two to three sentences of **more detailed background**, comprehensible  to scientists in related disciplines.

  One sentence clearly stating the **general problem** being addressed by  this particular study.

  One sentence summarizing the main result (with the words "**here we show**" or their equivalent).

  Two or three sentences explaining what the **main result** reveals in direct comparison to what was thought to be the case previously, or how the  main result adds to previous knowledge.

  One or two sentences to put the results into a more **general context**.

  Two or three sentences to provide a **broader perspective**, readily comprehensible to a scientist in any discipline.

  <!-- https://tinyurl.com/ybremelq -->

keywords          : "keywords"
wordcount         : "X"

bibliography      : ["bibliography.bib", "r-references.bib"]

floatsintext      : no
figurelist        : no
tablelist         : no
footnotelist      : no
linenumbers       : yes
mask              : no
draft             : no

documentclass     : "apa6"
classoption       : "man"
output            : papaja::apa6_pdf
---

```{r setup, include = FALSE}
library("papaja")
r_refs("r-references.bib")

```

```{r analysis-preferences}
# Seed for random number generation
set.seed(42)
knitr::opts_chunk$set(cache.extra = knitr::rand_seed)
```

@gomezDiffusionModelAccount2013 performed an experiment

# Methods
We report how we determined our sample size, all data exclusions (if any), all manipulations, and all measures in the study. <!-- 21-word solution (Simmons, Nelson & Simonsohn, 2012; retrieved from http://ssrn.com/abstract=2160588) -->

## Participants

## Material

## Procedure

## Data analysis
We used `r cite_r("r-references.bib")` for all our analyses.

# Results

# Discussion

\newpage

# References

\begingroup
\setlength{\parindent}{-0.5in}
\setlength{\leftskip}{0.5in}

<div id="refs" custom-style="Bibliography"></div>
\endgroup