### 2021-04-27 Lab meeting presentation April 2021 Results.Rmd

---
title: "Can we use masked priming in an online setting?"

subtitle: "The effect of prime exposure duration in online masked priming lexical decision"
author: "Bernhard Angele, Ana Baciero, Pablo Gomez, & Manuel Perea"
institute: "Bournemouth University, Universidad Nebrija, DePaul University, California State University San Bernardino, Universitat de Valencia"
date: "`r Sys.Date()`"
output:
  xaringan::moon_reader:
    css: [default, metropolis, metropolis-fonts] 
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---

# Doing research during a pandemic

- No eye-tracking!
- But lots of bored people sitting at home
- What are language-related paradigms that we can do remotely?
  - Surveys (not terribly interesting for many language researchers)
  - Self-paced reading
  - RSVP
  - Lexical decision tasks

---

# Is priming possible online?

- Overt priming certainly is
- But masked priming relies on very precise timing of the stimuli.
- Ken Forster did it a long time ago with GIFs:

  - 40 ms prime duration: 
  ![40 ms prime](http://www.u.arizona.edu/~kforster/priming/mp_demo_1.gif)
  - 100 ms prime duration: 
  ![100 ms prime](http://www.u.arizona.edu/~kforster/priming/mp_demo_2.gif)

- But a demo on a website and an online experiment are two very different things!

---

# Using the three Ps (Psychopy, Pavlovia, and Prolific) to collect data

- Psychopy is an experiment presentation software allowing precise display of stimuli
  - It's written in Python, but there is now the option of translating the experiments into Javascript (PsychoJS) for online presentation.
  - Pavlovia is a paid service that integrates with Psychopy and hosts the experiment code and the collected data
  - Prolific is a participant recruitment service (like Mechanical Turk, but more research-focused and less US-centric)

---

# Our study

- Based on Gomez, Perea, & Ratcliff (2013)

- Gomez et al. used the diffusion model to fit data from masked and unmasked prime presentation and found that masked and unmasked priming affect different model parameters:
  - masked identity priming affects the $T_{er}$ parameter which represents encoding processes
  - unmasked priming affects both $T_er$ and the drift rates parameter corresponding to the quality of information

- If we can show that the data from an online experiment pattern with the masked priming data from Gomez et al. rather than the unmasked priming data, that demonstrates the reliability of the online masked priming paradigm.

---

# Experiment 1

## Design

- Targets: 240 words and 240 nonwords
- Conditions: 
  - prime duration (33 vs. 50 ms)
    - this corresponds to 2 vs. 3 refresh cycles at 60 Hz
  - prime condition: identical vs. unrelated
- 60 word/nonword stimuli per group

## Participants

```{r load_exp1_data, message=FALSE, echo=FALSE, warning=FALSE}
library(xfun)
library(tidyverse)
library(sp)
library(rworldmap)

# Qualtrics survey data
exp1_all_participants <- read_csv("participant_data_exp1.csv")

# PsychoJS trial data
exp1 <- fs::dir_ls(path = "./data_exp1", glob = "*.csv") %>%
  map_dfr(read_csv, .id = "source", col_type = cols(
  .default = col_character(), rt = col_double(), corr = col_integer(), TrialID = col_integer())) %>% 
  filter(!is.na(TrialID) & TrialID < 1000) %>%
  select(source, participant, date, OS, frameRate, rt, corr, TrialID, StimulusType, Condition, PrimeDuration, Prime, Target)

exp1$device <- ifelse(exp1$OS %in% c("Linux armv7l", "Linux armv8l"), "android", "computer")

# only consider participants who have actually participated in the whole experiment
# also don't consider participants whose age is 99 (used for testing)
# one participant is duplicate. Remove the first occurrence (likely incompatible device)

exp1_actual_participants <- filter(exp1_all_participants, (PROLIFIC_PID %in% exp1$participant) & (nchar(PROLIFIC_PID) == 24) & !(PROLIFIC_PID %in% c("5fa3b4abbcfd0b6c243758bc")) & (`What is your age?` != 99) & (`Response ID` != "R_sU4qJ6UCe9jRKs9"))

# get participants with more than 80% correct responses

exp1_accuracy_by_participant <- exp1 %>% filter(participant %in% exp1_actual_participants$PROLIFIC_PID) %>% group_by(source) %>% summarise(acc = mean(corr == 1), N = n())

exp1_participants_to_include <- exp1_accuracy_by_participant %>% filter(N == 480 & acc >= .8)

exp1_participants_to_exclude <- exp1_accuracy_by_participant %>% filter(N != 480 | acc < .8)

#participant_list_for_prolific <- str_extract(string = participants_to_include$source, pattern = "[a-z0-9]{24}")

exp1_data_to_include <- exp1 %>% filter(source %in% exp1_participants_to_include$source & participant %in% exp1_actual_participants$PROLIFIC_PID) %>% mutate(StimulusType = StimulusType %>% factor(levels = c("NW", "WORD"), labels = c("Nonword", "Word")), Condition = Condition %>% factor(levels = c("ID", "UN"), labels = c("Identical", "Unrelated")), PrimeDuration = PrimeDuration %>% factor(levels = c(33, 50), labels = c("33 ms", "50 ms")), rt = rt * 1000)

# filter out any of the participants we eliminated from the Qualtrics file as well

exp1_actual_participants <- filter(exp1_actual_participants, (PROLIFIC_PID %in% exp1_data_to_include$participant))

exp1_participant_ages <- exp1_actual_participants$`What is your age?`

exp1_male_participants_nr <- filter(exp1_actual_participants, `What is your gender (optional)?` == "Male") %>% nrow

exp1_female_participants_nr <- filter(exp1_actual_participants, `What is your gender (optional)?` == "Female") %>% nrow

# there aren't actually any of these, but will leave the code in just in case
exp1_other_participants_nr <- filter(exp1_actual_participants, `What is your gender (optional)?` == "Other") %>% nrow

exp1_NA_participants_nr <- filter(exp1_actual_participants, is.na(`What is your gender (optional)?`)) %>% nrow

exp1_participant_coords <- exp1_actual_participants %>% select("Location Longitude", "Location Latitude")

#coordinates(exp1_actual_participants) <- c("Location Longitude","Location Latitude")

# from https://stackoverflow.com/questions/14334970/convert-latitude-and-longitude-coordinates-to-country-name-in-r

# The single argument to this function, points, is a data.frame in which:
#   - column 1 contains the longitude in degrees
#   - column 2 contains the latitude in degrees
coords2country = function(points)
{  
  countriesSP <- getMap(resolution='low')
  #countriesSP <- getMap(resolution='high') #you could use high res map from rworldxtra if you were concerned about detail

  # convert our list of points to a SpatialPoints object

  # pointsSP = SpatialPoints(points, proj4string=CRS(" +proj=longlat +ellps=WGS84 +datum=WGS84 +no_defs +towgs84=0,0,0"))

  #setting CRS directly to that from rworldmap
  pointsSP = SpatialPoints(points, proj4string=CRS(proj4string(countriesSP)))

  # use 'over' to get indices of the Polygons object containing each point 
  indices = over(pointsSP, countriesSP)

  # return the ADMIN names of each country
  indices$ADMIN  
  #indices$ISO3 # returns the ISO3 code 
  #indices$continent   # returns the continent (6 continent model)
  #indices$REGION   # returns the continent (7 continent model)
}

participant_coords <- exp1_actual_participants %>% select("Location Longitude", "Location Latitude")

exp1_actual_participants$country <- coords2country(participant_coords)

exp1_participant_countries <- table(exp1_actual_participants$country %>% as.character) #make into a character vector since the original output is a factor with as many levels as there are countries
```

## Participants

- `r exp1_all_participants$PROLIFIC_PID %>% unique %>% length` Prolific participants, with `r exp1_accuracy_by_participant %>% nrow` with `r exp1_actual_participants$PROLIFIC_PID %>% unique %>% length` completing the experiment.
- Participants (`r exp1_male_participants_nr` male, `r exp1_female_participants_nr` female) were aged from `r exp1_participant_ages %>% min` to `r exp1_participant_ages %>% max` (mean age: `r exp1_participant_ages %>% mean %>% round(2)`). Of the participants,  identified as male and `r exp1_female_participants_nr` identified as female. The experiment was only shown to participants who indicated that English was their first language in the Prolific screening questions. Based on their IP addresses, `r exp1_participant_countries["United Kingdom"] %>% ifelse(. < 11, numbers_to_words(.), .)` participants were based in the UK, `r exp1_participant_countries["United States of America"] %>% ifelse(. < 11, numbers_to_words(.), .)` were based in the US, `r exp1_participant_countries["Canada"] %>% ifelse(. < 11, numbers_to_words(.), .)` participants were based in Canada, and `r exp1_participant_countries["Ireland"] %>% ifelse(. < 11, numbers_to_words(.), .)` participants were based in Ireland. Two participants could not be localized in this way. Participants who completed the experiment were paid £1.25 for their participation (corresponding to £5/hour). All participants were naïve to the purpose of the experiment. Participants could use a desktop or laptop computer or a mobile device. Because of a technical display issue with PsychoJS and the Safari browser, participants who tried to access the experiment using that browser, including all participants on iOS devices, were advised to change browser or device and restart the experiment.

---
# Design

## Experiment 2

  - Like Experiment 1, but shorter prime duration condition (16 ms vs. 33 ms):
    - Targets: same 240 words and 240 nonwords
    - Conditions: 
      - prime duration (16 vs. 33 ms)
        - this corresponds to 1 vs. 2 refresh cycles at 60 Hz
      - prime condition: identical vs. unrelated
    - 60 word/nonword stimuli per group

---
# Results