### 2021-04-28 Lab meeting presentation April 2021.Rmd
Add raincloud plots

---
title: "Can we use masked priming in an online setting?"
subtitle: "The effect of prime exposure duration in online masked priming lexical decision"

author: "Bernhard Angele, Ana Baciero, Pablo Gomez, & Manuel Perea"

output:
  xaringan::moon_reader:
    css: [default, metropolis, metropolis-fonts] 
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---

```{r setup, echo=FALSE, message=FALSE}
library(tidyverse)
# for finding out where the participants were
library(xfun)
library(sp)
library(rworldmap)
# for the analyses
library(brms)
# for the plots
library(cowplot)
library(stargazer)
source("RainCloudPlots.R")
```

# Masked priming paradigm

-  One of the most important techniques to study early letter/word processes

![Trial diagram](MaskedPrimingDiagram.PNG)

<small>(Forster & Davis, 1984)

---

# Is masked priming possible online?

<br>

- Technology improvements + current situation have made online studies flourish

<br>

- Different paradigms are easily implemented in online settings:

  - Surveys
  - Self-paced reading
  - RSVP
  - Lexical decision tasks

<br>

- But masked priming relies on very precise timing of the stimuli

  - Issue: use of different devices to present the stimuli and record responses

---

# Here

<br>
<center>
<big>
**Is it possible to do masked priming experiments online?**

<br>

---

# But before diving into our study...

<br>

- A key phenomenon in laboratory masked priming lexical decision is that asked identity priming reflects a *savings effect*:
  - Lexical entry evaluation begins sooner for identity primes than for unrelated primes.

    >- RT distributions of the unrelated and identity pairs reflect a shift rather than a change in shape.

<br>

- Additionally, such shift should be similar in magnitude to the prime-target stimulus-onset asynchrony

---
# Model fits of observed data

<br>

- Gomez, Perea, & Ratcliff (2013) diffusion model fits:

  -  Masked identity priming affects the $T_{er}$ parameter (encoding processes)

  -  Unmasked priming affects both $T_{er}$ and the drift rates parameter (quality of information)

---

# Research question

<br>
<br>

Could we replicate these findings using an online setup?

<br>

- If we can show that the data from an online experiment pattern with the masked priming data from Gomez et al. rather than the unmasked priming data, that would show the reliability of the online masked priming paradigm.

---

# Popular online data collection method

<br>

- Integration of 3 pieces of software: *The three Ps*

  - **Psychopy**: experiment presentation software that allows precise display of stimuli

  - **Pavlovia**: paid service that hosts the experiment code and the collected data

  - **Prolific**: participant recruitment service (like Mechanical Turk, but more research-focused and less US-centric)

---

# The present study

<br>

- We used said data collection method to examine whether online masked priming studies follow the same pattern as in-lab masked priming studies, manipulating:

  - Prime condition:

    >- Identity vs. Unrelated

  - Prime exposure duration:

    >- 33.3 vs. 50 ms (Experiment 1)
    >- 16.6 vs. 33.3 ms (Experiment 2)

---

# Experiment 1

## Experimental Design

<br>

- Targets: 240 words and 240 nonwords

<br>

- Conditions:

  - prime duration (33 vs. 50 ms)
    >- Corresponding to 2 vs. 3 refresh cycles at 60 Hz

  - prime condition: identical vs. unrelated

<br>

- 60 word/nonword stimuli per group
---

# Experiment 1: Participants

```{r load_exp1_data, message=FALSE, echo=FALSE, warning=FALSE}

# Qualtrics survey data
exp1_all_participants <- read_csv("participant_data_exp1.csv")

# PsychoJS trial data
exp1 <- fs::dir_ls(path = "./data_exp1", glob = "*.csv") %>%
  map_dfr(read_csv, .id = "source", col_type = cols(
  .default = col_character(), rt = col_double(), corr = col_integer(), TrialID = col_integer())) %>% 
  filter(!is.na(TrialID) & TrialID < 1000) %>%
  select(source, participant, date, OS, frameRate, rt, corr, TrialID, StimulusType, Condition, PrimeDuration, Prime, Target)

exp1$device <- ifelse(exp1$OS %in% c("Linux armv7l", "Linux armv8l"), "android", "computer")

# only consider participants who have actually participated in the whole experiment
# also don't consider participants whose age is 99 (used for testing)
# one participant is duplicate. Remove the first occurrence (likely incompatible device)

exp1_actual_participants <- filter(exp1_all_participants, (PROLIFIC_PID %in% exp1$participant) & (nchar(PROLIFIC_PID) == 24) & !(PROLIFIC_PID %in% c("5fa3b4abbcfd0b6c243758bc")) & (`What is your age?` != 99) & (`Response ID` != "R_sU4qJ6UCe9jRKs9"))

# get participants with more than 80% correct responses

exp1_accuracy_by_participant <- exp1 %>% filter(participant %in% exp1_actual_participants$PROLIFIC_PID) %>% group_by(source) %>% summarise(acc = mean(corr == 1), N = n())

exp1_participants_to_include <- exp1_accuracy_by_participant %>% filter(N == 480 & acc >= .8)

exp1_participants_to_exclude <- exp1_accuracy_by_participant %>% filter(N != 480 | acc < .8)

#participant_list_for_prolific <- str_extract(string = participants_to_include$source, pattern = "[a-z0-9]{24}")

exp1_data_to_include <- exp1 %>% filter(source %in% exp1_participants_to_include$source & participant %in% exp1_actual_participants$PROLIFIC_PID) %>% mutate(StimulusType = StimulusType %>% factor(levels = c("NW", "WORD"), labels = c("Nonword", "Word")), Condition = Condition %>% factor(levels = c("ID", "UN"), labels = c("Identical", "Unrelated")), PrimeDuration = PrimeDuration %>% factor(levels = c(33, 50), labels = c("33 ms", "50 ms")), rt = rt * 1000)

# filter out any of the participants we eliminated from the Qualtrics file as well

exp1_actual_participants <- filter(exp1_actual_participants, (PROLIFIC_PID %in% exp1_data_to_include$participant))

exp1_participant_ages <- exp1_actual_participants$`What is your age?`

exp1_male_participants_nr <- filter(exp1_actual_participants, `What is your gender (optional)?` == "Male") %>% nrow

exp1_female_participants_nr <- filter(exp1_actual_participants, `What is your gender (optional)?` == "Female") %>% nrow

# there aren't actually any of these, but will leave the code in just in case
exp1_other_participants_nr <- filter(exp1_actual_participants, `What is your gender (optional)?` == "Other") %>% nrow

exp1_NA_participants_nr <- filter(exp1_actual_participants, is.na(`What is your gender (optional)?`)) %>% nrow

exp1_participant_coords <- exp1_actual_participants %>% select("Location Longitude", "Location Latitude")

#coordinates(exp1_actual_participants) <- c("Location Longitude","Location Latitude")

# from https://stackoverflow.com/questions/14334970/convert-latitude-and-longitude-coordinates-to-country-name-in-r

# The single argument to this function, points, is a data.frame in which:
#   - column 1 contains the longitude in degrees
#   - column 2 contains the latitude in degrees
coords2country = function(points)
{  
  countriesSP <- getMap(resolution='low')
  #countriesSP <- getMap(resolution='high') #you could use high res map from rworldxtra if you were concerned about detail

  # convert our list of points to a SpatialPoints object

  # pointsSP = SpatialPoints(points, proj4string=CRS(" +proj=longlat +ellps=WGS84 +datum=WGS84 +no_defs +towgs84=0,0,0"))

  #setting CRS directly to that from rworldmap
  pointsSP = SpatialPoints(points, proj4string=CRS(proj4string(countriesSP)))

  # use 'over' to get indices of the Polygons object containing each point 
  indices = over(pointsSP, countriesSP)

  # return the ADMIN names of each country
  indices$ADMIN  
  #indices$ISO3 # returns the ISO3 code 
  #indices$continent   # returns the continent (6 continent model)
  #indices$REGION   # returns the continent (7 continent model)
}

participant_coords <- exp1_actual_participants %>% select("Location Longitude", "Location Latitude")

exp1_actual_participants$country <- coords2country(participant_coords)

exp1_participant_countries <- table(exp1_actual_participants$country %>% as.character) #make into a character vector since the original output is a factor with as many levels as there are countries
```

- `r exp1_all_participants$PROLIFIC_PID %>% unique %>% length` Prolific participants, with `r exp1_actual_participants$PROLIFIC_PID %>% unique %>% length` actually completing the experiment with an accuracy of > .8.
- `r exp1_male_participants_nr` male and `r exp1_female_participants_nr` female participants 
- Aged from `r exp1_participant_ages %>% min` to `r exp1_participant_ages %>% max` (mean: `r exp1_participant_ages %>% mean %>% round(2)`). 
- Only shown to participants who indicated that English was their first language in the Prolific screening questions.
- Participant countries (according to Qualtrics and based on the IP Addresses, only including participants who could be localized)

```{r participant_location_table, as.is = TRUE, echo = FALSE}
knitr::kable(exp1_participant_countries %>% t, format = "html")

```
- Participants who completed the experiment were paid £1.25 for their participation (corresponding to £5/hour). 
- Participants could use a desktop or laptop computer or a mobile device.
  - But not Safari (issues with PsychoJS)

---
# Experiment 1: Results

```{r raincloud_exp1, echo=FALSE}
plot_raincloud(exp1_data_to_include %>% filter(corr == 1 & rt < 1500), x_column = "Condition", y_column = "rt", x_label = "Condition", y_label = "RT in ms (correct responses)", main_title = "RT by condition and word type") + 
  # put labels on the mean for easier comparison
  stat_summary(fun.data = function(x) {data.frame(label = format(mean(x), digits = 2), y = mean(x))}, 
                                      geom = "text",
                                      #width = 1,
                                      position = position_nudge(x = 0.5, y = 0),
                                      #size = 0.2,
                                      color = "black") +
  facet_grid(cols = vars(PrimeDuration), rows = vars(StimulusType), labeller = label_context)
```

---
# Experiment 1: EDA
<br>

![Trial diagram](DeltaE1_PrimingEffect.png)

---
# Experiment 2

## Experimental Design

<br>

- Targets: 240 words and 240 nonwords

<br>

- Conditions:

  - prime duration (16 vs. 33.3 ms)
    >- Corresponding to 1 vs. 2 refresh cycles at 60 Hz

  - prime condition: identical vs. unrelated

<br>

- 60 word/nonword stimuli per group

---
# Experiment 2: Participants

```{r load-exp2-data, as.is = TRUE,echo=FALSE, warning=FALSE}
#load raw Exp2 data

exp2_all_participants <- read_csv("participant_data_exp2.csv")

exp2 <- fs::dir_ls(path = "./data_exp2", glob = "*.csv") %>%
  map_dfr(read_csv, .id = "source", col_type = cols(
  .default = col_character(), rt = col_double(), corr = col_integer(), TrialID = col_integer())) %>% 
  filter(!is.na(TrialID) & TrialID < 1000) %>%
  select(source, participant, date, OS, frameRate, rt, corr, TrialID, StimulusType, Condition, PrimeDuration, Prime, Target)

exp2$device <- ifelse(exp2$OS %in% c("Linux armv7l", "Linux armv8l"), "android", "computer")

# only consider participants who have actually participated in the whole experiment
# also don't consider participants whose age is 99 (used for testing)
# one participant is duplicate. Remove the first occurrence (likely incompatible device)

exp2_actual_participants <- filter(exp2_all_participants, (PROLIFIC_PID %in% exp2$participant) & (nchar(PROLIFIC_PID) == 24) & !(PROLIFIC_PID %in% c("5fa3b4abbcfd0b6c243758bc")) & (`What is your age?` != 99) & !(`Response ID` %in% c("R_2c2dVRNmGvI6CH9", "R_3ELutaT6GNe5mmR", "R_XX7jgUk2NbIi41j", "R_125QKzJfHwoGHaT")))

# get participants with more than 80% correct responses

exp2_accuracy_by_participant <- exp2 %>% filter(participant %in% exp2_actual_participants$PROLIFIC_PID) %>% group_by(source) %>% summarise(acc = mean(corr == 1), N = n())

exp2_participants_to_include <- exp2_accuracy_by_participant %>% filter(N == 480 & acc >= .8)

exp2_participants_to_exclude <- exp2_accuracy_by_participant %>% filter(N != 480 | acc < .8)

#participant_list_for_prolific <- str_extract(string = participants_to_include$source, pattern = "[a-z0-9]{24}")

exp2_data_to_include <- exp2 %>% filter(source %in% exp2_participants_to_include$source & participant %in% exp2_actual_participants$PROLIFIC_PID) %>% mutate(StimulusType = StimulusType %>% factor(levels = c("NW", "WORD"), labels = c("Nonword", "Word")), Condition = Condition %>% factor(levels = c("ID", "UN"), labels = c("Identical", "Unrelated")), PrimeDuration = PrimeDuration %>% factor(levels = c(16, 33), labels = c("16 ms", "33 ms")), rt = rt * 1000)

# filter out any of the participants we eliminated from the Qualtrics file as well

exp2_actual_participants <- filter(exp2_actual_participants, (PROLIFIC_PID %in% exp2_data_to_include$participant))

exp2_participant_ages <- exp2_actual_participants$`What is your age?`

exp2_male_participants_nr <- filter(exp2_actual_participants, `What is your gender (optional)?` == "Male") %>% nrow

exp2_female_participants_nr <- filter(exp2_actual_participants, `What is your gender (optional)?` == "Female") %>% nrow

# there aren't actually any of these, but will leave the code in just in case
exp2_other_participants_nr <- filter(exp2_actual_participants, `What is your gender (optional)?` == "Other") %>% nrow

exp2_NA_participants_nr <- filter(exp2_actual_participants, is.na(`What is your gender (optional)?`)) %>% nrow

exp2_participant_coords <- exp2_actual_participants %>% select("Location Longitude", "Location Latitude")

# coords2country function definition in exp1 chunk above
exp2_actual_participants$country <- coords2country(exp2_participant_coords)

exp2_participant_countries <- table(exp2_actual_participants$country %>% as.character) #make into a character vector since the original output is a factor with as many levels as there are countries

#paste0(participant_list_for_prolific, sep = "\n") %>% writeClipboard() # works on windows

exclude_list_for_prolific <- str_extract(string = exp2_participants_to_exclude$source, pattern = "[a-z0-9]{24}") %>% paste0(collapse = "\n")

```

- `r exp2_all_participants$PROLIFIC_PID %>% unique %>% length` Prolific participants, with `r exp2_actual_participants$PROLIFIC_PID %>% unique %>% length` actually completing the experiment with an accuracy of > .8.
- `r exp2_male_participants_nr` male and `r exp2_female_participants_nr` female participants 
- Aged from `r exp2_participant_ages %>% min` to `r exp2_participant_ages %>% max` (mean: `r exp2_participant_ages %>% mean %>% round(2)`). 
- Only shown to participants who indicated that English was their first language in the Prolific screening questions.
- Participant countries (according to Qualtrics and based on the IP Addresses, only including participants who could be localized)

```{r participant_location_table_exp2, as.is = TRUE, echo = FALSE}
knitr::kable(exp2_participant_countries %>% t, format = "html")

```

---
# Experiment 2: Results

```{r raincloud_exp2, echo=FALSE}
plot_raincloud(exp2_data_to_include %>% filter(corr == 1 & rt < 1500), x_column = "Condition", y_column = "rt", x_label = "Condition", y_label = "RT in ms (correct responses)", main_title = "RT by condition and word type") + 
  # put labels on the mean for easier comparison
  stat_summary(fun.data = function(x) {data.frame(label = format(mean(x), digits = 2), y = mean(x))}, 
                                      geom = "text",
                                      #width = 1,
                                      position = position_nudge(x = 0.5, y = 0),
                                      #size = 0.2,
                                      color = "black") +
  facet_grid(cols = vars(PrimeDuration), rows = vars(StimulusType), labeller = label_context)
```

---
# Experiment 2: EDA
<br>

![Trial diagram](DeltaE2_PrimingEffect.png)
---
# Inferentials
<br>

---
# Conclusions
<br>