### 2021-09-21 Appendix_test.Rmd

---
appendix: 
  - "./Appendix_A.Rmd"
  - "./Appendix_B.Rmd"
output: papaja::apa6_pdf
---

This is the manuscript.

```{r setup, include = FALSE}
library(tidyverse)
library("see")
library("papaja")
library(patchwork)
library(xfun)
library(sp)
library(rworldmap)
library("brms")
library("bayestestR")

r_refs("r-references.bib")

```

```{r load_trial_data, message=FALSE, echo=FALSE}
library(readxl)

trials <- read_xlsx(path = "html/resources/trials_list0.xlsx")

word_statistics <- read_xlsx(path = "online_MaskedPriming_materials_v1.xlsx")

word_statistics$Word_upper <- word_statistics$Word %>% toupper()

trials_word_stats <- left_join(trials, word_statistics, by = c("Target" = "Word_upper"))

```