### 2021-11-06 previous revisions/EDA_OnlineMaskedPriming.Rmd

---
title: "EDA"
author: "Ana Baciero"
date: "20/4/2021"
output: pdf_document
---

```{r setup, include=FALSE}

library(tidyverse)
library(ggplot2)
library(papaja)
library(patchwork)

```

# Data sets:

## Experiment 1

```{r Experiment 1, include=FALSE}

# Qualtrics survey data
exp1_all_participants <- read_csv("participant_data_exp1.csv")

# PsychoJS trial data
exp1 <- fs::dir_ls(path = "./data_exp1", glob = "*.csv") %>%
  map_dfr(read_csv, .id = "source", col_type = cols(
    .default = col_character(), rt = col_double(), corr = col_integer(), TrialID = col_integer())) %>% 
  filter(!is.na(TrialID) & TrialID < 1000) %>%
  select(source, participant, date, OS, frameRate, rt, corr, TrialID, StimulusType, Condition, PrimeDuration, Prime, Target)

exp1$device <- ifelse(exp1$OS %in% c("Linux armv7l", "Linux armv8l"), "android", "computer")

# only consider participants who have actually participated in the whole experiment
# also don't consider participants whose age is 99 (used for testing)
# one participant is duplicate. Remove the first occurrence (likely incompatible device)

exp1_actual_participants <- filter(exp1_all_participants, (PROLIFIC_PID %in% exp1$participant) & (nchar(PROLIFIC_PID) == 24) & !(PROLIFIC_PID %in% c("5fa3b4abbcfd0b6c243758bc")) & (`What is your age?` != 99) & (`Response ID` != "R_sU4qJ6UCe9jRKs9"))

# get participants with more than 80% correct responses

exp1_accuracy_by_participant <- exp1 %>% filter(participant %in% exp1_actual_participants$PROLIFIC_PID) %>% group_by(source) %>% summarise(acc = mean(corr == 1), N = n())

exp1_participants_to_include <- exp1_accuracy_by_participant %>% filter(N == 480 & acc >= .8)

exp1_participants_to_exclude <- exp1_accuracy_by_participant %>% filter(N != 480 | acc < .8)

#participant_list_for_prolific <- str_extract(string = participants_to_include$source, pattern = "[a-z0-9]{24}")

exp1_data_to_include <- exp1 %>% 
  filter(source %in% exp1_participants_to_include$source & participant %in% exp1_actual_participants$PROLIFIC_PID) %>% 
  mutate(StimulusType = StimulusType %>% 
           factor(levels = c("NW", "WORD"), labels = c("Nonword", "Word")), Condition = Condition %>% 
           factor(levels = c("ID", "UN"), labels = c("Identical", "Unrelated")), PrimeDuration = PrimeDuration %>% factor(levels = c(33, 50), labels = c("33 ms", "50 ms")), rt = rt * 1000) %>% 
  filter(rt > 250 & rt < 2000)
```

```{r E1 Head, tidy=TRUE}

E1<- exp1_data_to_include%>% 
  filter(rt > 250 & rt < 2000)%>%
  select(participant, TrialID, StimulusType, Condition, PrimeDuration, Prime, Target, corr, rt)%>%
  mutate(participant=factor(participant),
         TrialID = factor(TrialID))

head(E1)
```

## Experiment 2

```{r Experiment 2, include=FALSE}

exp2_all_participants <- read_csv("participant_data_exp2.csv")

exp2 <- fs::dir_ls(path = "./data_exp2", glob = "*.csv") %>%
  map_dfr(read_csv, .id = "source", col_type = cols(
    .default = col_character(), rt = col_double(), corr = col_integer(), TrialID = col_integer())) %>% 
  filter(!is.na(TrialID) & TrialID < 1000) %>%
  select(source, participant, date, OS, frameRate, rt, corr, TrialID, StimulusType, Condition, PrimeDuration, Prime, Target)

exp2$device <- ifelse(exp2$OS %in% c("Linux armv7l", "Linux armv8l"), "android", "computer")

# only consider participants who have actually participated in the whole experiment
# also don't consider participants whose age is 99 (used for testing)
# one participant is duplicate. Remove the first occurrence (likely incompatible device)

exp2_actual_participants <- filter(exp2_all_participants, (PROLIFIC_PID %in% exp2$participant) & (nchar(PROLIFIC_PID) == 24) & !(PROLIFIC_PID %in% c("5fa3b4abbcfd0b6c243758bc")) & (`What is your age?` != 99) & !(`Response ID` %in% c("R_2c2dVRNmGvI6CH9", "R_3ELutaT6GNe5mmR", "R_XX7jgUk2NbIi41j", "R_125QKzJfHwoGHaT")))

# get participants with more than 80% correct responses

exp2_accuracy_by_participant <- exp2 %>% filter(participant %in% exp2_actual_participants$PROLIFIC_PID) %>% group_by(source) %>% summarise(acc = mean(corr == 1), N = n())

exp2_participants_to_include <- exp2_accuracy_by_participant %>% filter(N == 480 & acc >= .8)

exp2_participants_to_exclude <- exp2_accuracy_by_participant %>% filter(N != 480 | acc < .8)

#participant_list_for_prolific <- str_extract(string = participants_to_include$source, pattern = "[a-z0-9]{24}")

exp2_data_to_include <- exp2 %>% filter(source %in% exp2_participants_to_include$source & participant %in% exp2_actual_participants$PROLIFIC_PID) %>% mutate(StimulusType = StimulusType %>% factor(levels = c("NW", "WORD"), labels = c("Nonword", "Word")), Condition = Condition %>% factor(levels = c("ID", "UN"), labels = c("Identical", "Unrelated")), PrimeDuration = PrimeDuration %>% factor(levels = c(16, 33), labels = c("16 ms", "33 ms")), rt = rt * 1000) %>% filter(rt > 250 & rt < 2000)

```

```{r E2 Head, tidy=TRUE}

E2<- exp2_data_to_include%>%
  select(participant, TrialID, StimulusType, Condition, PrimeDuration, Prime, Target, corr, rt)%>%
  mutate(participant=factor(participant),
         TrialID = factor(TrialID))

head(E2)

```

\newpage

# Delta Plots

```{r Deltas E1, message=FALSE, warning=FALSE}

quibble <- function(x, q = seq(.1, .9, .2)) {
  tibble(x = quantile(x, q), q = q)
}

E1 %>%
  group_by(participant, StimulusType, Condition, PrimeDuration,corr) %>%
  summarise(RT = list(quibble(rt, seq(.1, .9, .2)))) %>% 
  tidyr::unnest(RT) -> data.E1.quantiles

data.E1.quantiles %>%
  filter(corr==1) %>%
  select(-corr) %>%
  group_by(StimulusType,PrimeDuration,Condition,q) %>%
  summarize(RT=mean(x))%>%
  ungroup()-> vincentiles.E1

#Priming Effect (unrelated - identity):

vincentiles.E1 %>%
  arrange(desc(Condition))%>%
  group_by(StimulusType,PrimeDuration,q) %>%
  summarize(MRT=mean(RT),
            Delta=diff(rev(RT)))%>%
  ungroup()%>%
ggplot(aes(y=Delta, x=MRT)) +
  geom_line()+
  geom_point()+
  ylim(-60, +60)+
  xlim(450, 950)+
  xlab("Mean RT per quantile")+
  ylab("Priming effect (delta)")+
  ggtitle("Experiment 1")+
  facet_wrap(~StimulusType + PrimeDuration)+
  theme_minimal()-> DeltaE1.PrimingEffect

DeltaE1.PrimingEffect

#Prime Duration Effect (33ms - 50 ms):

vincentiles.E1 %>%
  group_by(StimulusType,Condition,q) %>%
 summarize(MRT=mean(RT),
            Delta=diff(rev(RT)))%>%
ggplot(aes(y=Delta, x=MRT)) +
  geom_line()+
  geom_point()+
  ylim(-60, +60)+
  xlim(450, 950)+
  xlab("Mean RT per quantile")+
  ylab("Prime duration effect (delta)")+
  ggtitle("Experiment 1")+
  facet_wrap(~StimulusType + Condition)+
  theme_minimal()-> DeltaE1.PrimeDurationEffect

DeltaE1.PrimeDurationEffect

```

```{r Deltas E1 Words, message=FALSE, warning=FALSE}

#Priming Effect (unrelated - identity):

vincentiles.E1 %>%
  filter(StimulusType == "Word")%>%
  arrange(desc(Condition))%>%
  group_by(PrimeDuration,q) %>%
  summarize(MRT=mean(RT),
            Delta=diff(rev(RT)))%>%
  ungroup()%>%
ggplot(aes(y=Delta, x=MRT)) +
  geom_line(size = .75)+
  geom_point(size = 2)+
  geom_hline(yintercept = 0, color = "darkmagenta", alpha = .25)+
  ylim(-60, +60)+
  xlim(450, 850)+
  xlab("Mean RT per quantile")+
  ylab("Effect (in ms)")+
  ggtitle("Priming Effect")+
  facet_wrap(~PrimeDuration)+
  theme_minimal()+
  theme(plot.title = element_text(size = 12),
        axis.text=element_text(size=10),
        axis.title=element_text(size=10))-> DeltaE1.PrimingEffect.W

#Prime Duration Effect (33ms - 50 ms):

vincentiles.E1 %>%
  filter(StimulusType == "Word")%>%
  group_by(Condition,q) %>%
  summarize(MRT=mean(RT),
            Delta=diff(rev(RT)))%>%
  ggplot(aes(y=Delta, x=MRT)) +
  geom_line(size = .75)+
  geom_point(size = 2)+
  geom_hline(yintercept = 0, color = "darkmagenta", alpha = .25)+
  ylim(-60, +60)+
  xlim(450, 850)+
  xlab("Mean RT per quantile")+
  ylab("Effect (in ms)")+
  ggtitle("Prime duration effect")+
  facet_wrap(~ Condition)+
  theme_minimal()+
  theme(plot.title = element_text(size = 12),
        axis.text=element_text(size=10),
        axis.title=element_text(size=10))-> DeltaE1.PrimeDurationEffect.W

#pdf("Fig_deltasE1.pdf", height = 5)

cowplot::plot_grid(DeltaE1.PrimingEffect.W,
                   DeltaE1.PrimeDurationEffect.W,
                   ncol = 1,
                   labels = c("A", "B"))+
  plot_annotation(title = "Experiment 1: 33 vs. 50 ms",
                  theme = theme(plot.title = element_text(size = 14)))

#dev.off()

```

```{r Deltas E2, message=FALSE, warning=FALSE}

E2 %>%
  group_by(participant, StimulusType, Condition, PrimeDuration,corr) %>%
  summarise(RT = list(quibble(rt, seq(.1, .9, .2)))) %>% 
  tidyr::unnest(RT) -> data.E2.quantiles

data.E2.quantiles %>%
  filter(corr==1) %>%
  select(-corr) %>%
  group_by(StimulusType,PrimeDuration,Condition,q) %>%
  summarize(RT=mean(x))%>%
  ungroup()-> vincentiles.E2

#Priming Effect (unrelated - identity):

vincentiles.E2 %>%
  arrange(desc(Condition))%>%
  group_by(StimulusType,PrimeDuration,q) %>%
  summarize(MRT=mean(RT),
            Delta=diff(rev(RT)))%>%
  ungroup()%>%
ggplot(aes(y=Delta, x=MRT)) +
  geom_line()+
  geom_point()+
  ylim(-60, +60)+
  xlim(450, 950)+
  xlab("Mean RT per quantile")+
  ylab("Priming effect (delta)")+
  ggtitle("Experiment 2")+
  facet_wrap(~StimulusType + PrimeDuration)+
  theme_minimal() -> DeltaE2.PrimingEffect

DeltaE2.PrimingEffect

#Prime Duration Effect (16ms - 33 ms):

vincentiles.E2 %>%
  group_by(StimulusType,Condition,q) %>%
 summarize(MRT=mean(RT),
            Delta=diff(rev(RT)))%>%
ggplot(aes(y=Delta, x=MRT)) +
  geom_line()+
  geom_point()+
  ylim(-60, +60)+
  xlim(450, 950)+
  xlab("Mean RT per quantile")+
  ylab("Prime duration effect (delta)")+
  ggtitle("Experiment 2")+
  facet_wrap(~StimulusType + Condition)+
  theme_minimal() -> DeltaE2.PrimeDurationEffect

DeltaE2.PrimeDurationEffect

```

```{r Deltas E2 Words, message=FALSE, warning=FALSE}

#Priming Effect (unrelated - identity):

vincentiles.E2 %>%
  filter(StimulusType == "Word")%>%
  arrange(desc(Condition))%>%
  group_by(PrimeDuration,q) %>%
  summarize(MRT=mean(RT),
            Delta=diff(rev(RT)))%>%
  ungroup()%>%
ggplot(aes(y=Delta, x=MRT)) +
  geom_line(size = .75)+
  geom_point(size = 2)+
  geom_hline(yintercept = 0, color = "darkmagenta", alpha = .25)+
  ylim(-60, +60)+
  xlim(450, 850)+
  xlab("Mean RT per quantile")+
  ylab("Effect (in ms)")+
  ggtitle("Priming Effect")+
  facet_wrap(~PrimeDuration)+
  theme_minimal()+
  theme(plot.title = element_text(size = 12),
        axis.text=element_text(size=10),
        axis.title=element_text(size=10))-> DeltaE2.PrimingEffect.W

#Prime Duration Effect (16ms - 33 ms):

vincentiles.E2 %>%
  filter(StimulusType == "Word")%>%
  group_by(Condition,q) %>%
  summarize(MRT=mean(RT),
            Delta=diff(rev(RT)))%>%
  ggplot(aes(y=Delta, x=MRT)) +
  geom_line(size = .75)+
  geom_point(size = 2)+
  geom_hline(yintercept = 0, color = "darkmagenta", alpha = .25)+
  ylim(-60, +60)+
  xlim(450, 850)+
  xlab("Mean RT per quantile")+
  ylab("Effect (in ms)")+
  ggtitle("Prime duration effect")+
  facet_wrap(~ Condition)+
  theme_minimal()+
  theme(plot.title = element_text(size = 12),
        axis.text=element_text(size=10),
        axis.title=element_text(size=10))-> DeltaE2.PrimeDurationEffect.W

#pdf("Fig_deltasE2.pdf", height = 5)

cowplot::plot_grid(DeltaE2.PrimingEffect.W,
                   DeltaE2.PrimeDurationEffect.W,
                   ncol = 1,
                   labels = c("A", "B"))+
  plot_annotation(title = "Experiment 2: 16 vs. 33 ms",
                  theme = theme(plot.title = element_text(size = 14)))

```

\newpage

# CAF

```{r Experiment 1 CAF, message=FALSE, warning=FALSE}

E1 %>%
  group_by(participant, StimulusType, Condition, PrimeDuration) %>%
  mutate(Quantile = cut_number(rt, n = 10, labels = c(1,1,2,2,3,3,4,4,5,5))) %>%
  ungroup() %>%
  group_by(participant, StimulusType, Condition, PrimeDuration, Quantile) %>%
  summarize(Macc=mean(corr),
            Mrt=mean(rt)) %>%
  ungroup() -> data.caf.E1

data.caf.E1 %>%
  filter(StimulusType=="Word")%>%
  group_by(Condition, PrimeDuration, Quantile) %>%
    summarize(Macc2=mean(Macc),
            Mrt2=mean(Mrt)) %>%
  ggplot(aes(x=Mrt2, y=Macc2, )) +
geom_line(alpha=.3, size=1.2) +
  geom_point()+
  ylim(0.5, 1)+
  xlim(450, 1000)+
  facet_wrap(~PrimeDuration+Condition)+
  ggtitle("Accuracy as a function of RT in Experiment 1. Word data")+
  theme_minimal()+
  xlab("RT")+
  ylab("Accuracy") -> CAF.E1.W

CAF.E1.W

data.caf.E1 %>%
  filter(StimulusType=="Nonword")%>%
  group_by(Condition, PrimeDuration, Quantile) %>%
    summarize(Macc2=mean(Macc),
            Mrt2=mean(Mrt)) %>%
  ggplot(aes(x=Mrt2, y=Macc2, )) +
  geom_line(alpha=.3, size=1.2) +
  geom_point()+
  ylim(0.5, 1)+
  xlim(450, 1000)+
  facet_wrap(~PrimeDuration+Condition)+
  ggtitle("Accuracy as a function of RT in Experiment 1. Nonword data")+
  theme_minimal()+
  xlab("RT")+
  ylab("Accuracy") -> CAF.E1.NW

CAF.E1.NW

```

```{r Experiment 2 CAF, message=FALSE, warning=FALSE}

E2 %>%
  group_by(participant, StimulusType, Condition, PrimeDuration) %>%
  mutate(Quantile = cut_number(rt, n = 10, labels = c(1,1,2,2,3,3,4,4,5,5))) %>%
  ungroup() %>%
  group_by(participant, StimulusType, Condition, PrimeDuration, Quantile) %>%
  summarize(Macc=mean(corr),
            Mrt=mean(rt)) %>%
  ungroup() -> data.caf.E2

data.caf.E2 %>%
  filter(StimulusType=="Word")%>%
  group_by(Condition, PrimeDuration, Quantile) %>%
    summarize(Macc2=mean(Macc),
            Mrt2=mean(Mrt)) %>%
  ggplot(aes(x=Mrt2, y=Macc2, )) +
geom_line(alpha=.3, size=1.2) +
  geom_point()+
  ylim(0.5, 1)+
  xlim(450, 1000)+
  facet_wrap(~PrimeDuration+Condition)+
  ggtitle("Accuracy as a function of RT in Experiment 2. Word data")+
  theme_minimal()+
  xlab("RT")+
  ylab("Accuracy") -> CAF.E2.W

CAF.E2.W

data.caf.E2 %>%
  filter(StimulusType=="Nonword")%>%
  group_by(Condition, PrimeDuration, Quantile) %>%
    summarize(Macc2=mean(Macc),
            Mrt2=mean(Mrt)) %>%
  ggplot(aes(x=Mrt2, y=Macc2, )) +
  geom_line(alpha=.3, size=1.2) +
  geom_point()+
  ylim(0.5, 1)+
  xlim(450, 1000)+
  facet_wrap(~PrimeDuration+Condition)+
  ggtitle("Accuracy as a function of RT in Experiment 2. Nonword data")+
  theme_minimal()+
  xlab("RT")+
  ylab("Accuracy") -> CAF.E2.NW

CAF.E2.NW

```